---
title: 'Study title'
author: "your name"
date: "date"
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T)
solution <- T
```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

#Introduction:


## I. Data management
Data mangement is essential to the rest of the analytical process. 

The learning objectives are:
- load or read your dataset in R
- verify the dimension of your dataset
- verify the type of your variables and convert them if needed

1. Read the data in R

```{r load data}

```

2. Count the statistical units and the variables
```{r dimensions}

```

3. List the variables (using their names) related to the patients characteristics, the outcome, and the risk factors or potential counfounders.

4. Verify that the variable types (quantitative, qualitative) are correct, if not convert them.

```{r dimensions2}

```

## II. Descriptive statistics
Learning objectives: Describe the study population using the most appropriate indicators

1. Centrality measures: 
  a. Compute the mean, median, first and third quartile of each relevant variables.
  b. Do the same thing but grouped by the outcome variable (using by).
```{r descStat1}

```
2. Dispersion measures: 
  a. Compute the range, variance and standard deviation of each relevant variables. 
  b. Do the same thing but grouped by the outcome variable (using by).
```{r descStat2}

```
3. Frequency measures: 
  a. What is the purpose of the tabulate() function ? Use it on the age variable and interpret the result.
  b. What is the purpose of the which.max() function ? Use it on the previous result and interpret.
  c. Use the table function on two relevant variables you want to cross. Interpret. 
```{r descStat3}

```

## III. Visualization
Learning objectives: Describe the study population using the most appropriate visualization tools.
1. Build at least one histogram, bar chart, pie chart, and box-plot
  a. Find the relevant variables to visualize with each plot.
  b. Customize each title, axis labels and legends.
```{r plot1}

```
2. Using "layout()", find a relevant comparison to highlight by ploting the selected grouped data side by side.
```{r plot2}

```


## IV. Statistical tests (Ref Module 111 - Day 11 September 28. Homework 2)
Formulate statistical hypothesis according to the objective aimed by the study
Apply the statistical test using the R or STATA software and to interpret the results
Measure the strength of the association between two quantitative or qualitative variables and
interpret it
```{r tests}

```

## V. Linear regression (Ref. Module 106 - Day 16, November 9. Homework 3)
Learning objectives: Interpret a scatter plot
Measure the strength of the association between two quantitative or qualitative variables and
interpret it
Build a linear model
Follow the step by step procedure to obtain the best model and interpret it
Assess the validity of a linear model


```{r scatterplot}

```

