---
title: 'Study title'
author: "your name"
date: "date"
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T)
solution <- T
```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

#Introduction:


## I. Data management
Data mangement is essential to the rest of the analytical process. 

The learning objectives are:
- load or read your dataset in R
- verify the dimension of your dataset
- verify the type of your variables and convert them if neeeded

1. Read the data in R

```{r load data}

```

2. Count the statistical units and the variable
```{r dimensions}

```

3. List the variables (using their names) that related to the patients characteristics, the outcome, and the risk factors or potential counfounders.

4. Verify that the variable type (quantitative, qualitative) are correct, if not convert it.

```{r dimensions}

```

## II. Descriptive statistics
Learning objectives: Describe the study population using the most appropriate indicators
```{r desc stat}

```

## III. Visualization
Learning objectives: Describe the study population using the most appropriate indicators

```{r plot}

```

## IV. Statistical tests (Ref Modeul 111 - Day 11 September 28. Homework 2)
Formulate statistical hypothesis according to the objective aimed by the study
Apply the statistical test using the R or STATA software and to interpret the results
Measure the strength of the association between two quantitative or qualitative variables and
interpret it
```{r tests}

```

## V. Linear regression (Ref. Module 106 - Day 16, November 9. Homework 3)
Learning objectives: Interpret a scatter plot
Measure the strength of the association between two quantitative or qualitative variables and
interpret it
Build a linear model
Follow the step by step procedure to obtain the best model and interpret it
Assess the validity of a linear model


```{r scatterplot}

```

