---
title: 'Study title'
author: "your name"
date: "date"
output: pdf_document
---

#Introduction:
Every data analysis goes through most the steps listed below. Using the template below start your own analysis of the Scabies or the Binge drinking datasets available on REAL. 

## I. Data management
Learning objectives: Prepare the data for statistical analysis

1. Load data store in a csv file in R
```{r load data}
scabies=read.csv("ScabiesData.csv", sep=";", header=T)
```

2. Verify the dimension of your dataset by counting the statistical units and the variables
```{r dimensions}
nrow(scabies)
ncol(scabies)
```

3. List the variables (using their names) related to the patients' characteristics, the outcome, and the risk factors or potential counfounders.
```{r variable names}
names(scabies)
```

4. Verify that the variable types (quantitative, qualitative) are correct, if not convert them.
```{r variable type}
summary(scabies)
str(scabies)
scabies$SEXE = factor(scabies$SEXE, levels=c(1,2), labels=c("Male", "Female"))
scabies$STATUS = factor(scabies$STATUS, levels=c(1,2), labels=c("Patient", "Staff"))
scabies$SCABIES = factor(scabies$SCABIES, levels=c(1,0), labels=c("Yes", "No"))
scabies$START = as.Date(scabies$START, "%d/%m/%Y")
scabies$PRURIT = factor(scabies$PRURIT, levels=c(1,2), labels=c("Yes", "No"))
scabies$SILLON = factor(scabies$SILLON, levels=c(1,2), labels=c("Yes", "No"))
scabies$SARCOPTE = factor(scabies$SARCOPTE, levels=c(1,2), labels=c("Yes", "No"))
scabies$LOCAL = factor(scabies$LOCAL, levels=c(1,2,3,7), labels=c("Hands", "Members","Trunk","Everywhere"))
scabies$ASCABIOL = factor(scabies$ASCABIOL, levels=c(1,2), labels=c("Yes", "No"))
scabies$FLOOR = factor(scabies$FLOOR, levels=c(1,2,3), labels=c("First", "Second","Third"))
scabies$NEIGHBOR = factor(scabies$NEIGHBOR, levels=c(1,2), labels=c("Yes", "No"))
scabies$STATE = factor(scabies$STATE, levels=c(1,2,3), labels=c("Valid", "Limited","Bed"))
scabies$MEALS = factor(scabies$MEALS, levels=c(1,2), labels=c("DiningRoom", "Bedroom"))
scabies$MOVE = factor(scabies$MOVE, levels=c(1,2), labels=c("Yes", "No"))
scabies$VISITS = factor(scabies$VISITS, levels=c(1,2,3,4), labels=c("Frequent", "Sporadic","Rare","None"))
scabies$OUTSIDE = factor(scabies$OUTSIDE, levels=c(1,2,3,4), labels=c("Frequent", "Sporadically","Rarely","Never"))
scabies$CONTACT = factor(scabies$CONTACT, levels=c(1,2), labels=c("Yes", "No"))
scabies$WORKSHIFT = factor(scabies$WORKSHIFT, levels=c(1,2), labels=c("Day", "Night"))
```


## II. Descriptive statistics
Learning objectives: Describe the study population using the most appropriate indicators

1. Centrality measures: 
  a. Compute the mean, median, first and third quartile of each relevant variables.
  b. Do the same thing but grouped by the outcome variable (using by).
```{r descStat1}
summary(scabies$AGE)
by(scabies$AGE,scabies$SCABIES, summary)
```
2. Dispersion measures: 
  a. Compute the range, variance and standard deviation of each relevant variables. 
  b. Do the same thing but grouped by the outcome variable (using by).
```{r descStat2}
range(scabies$AGE)
var(scabies$AGE)
sd(scabies$AGE)
by(scabies$AGE,scabies$SCABIES,range)
by(scabies$AGE,scabies$SCABIES,var)
by(scabies$AGE,scabies$SCABIES,sd)
```
3. Frequency measures: 
  a. What is the purpose of the tabulate() function ? Use it on the age variable and interpret the result.
  b. What is the purpose of the which.max() function ? Use it on the previous result and interpret.
  c. Use the table and prop.table functions on two relevant variables you want to cross. Interpret. 
```{r descStat3}
tabulate(scabies$AGE)
which.max(tabulate(scabies$AGE))
table(scabies$SEXE,scabies$SCABIES)
```

## III. Visualization
Learning objectives: Describe the study population using the most appropriate visualization tools.
1. Build at least one histogram, bar chart, pie chart, and box-plot
  a. Find the relevant variables to visualize with each plot.
  b. Customize each title, axis labels and legends.
```{r plot1}
hist(scabies$AGE, main="Distribution of ages", xlab="years")
hist(scabies$START, "months", main="Scabies epidemic curve", xlab="months")
boxplot(AGE ~ SCABIES, data=scabies, main="Distribution of ages by scabies status", xlab="Scabies", ylab="age (years)")
barplot(table(scabies$SCABIES, scabies$STATE))
```
2. Using "layout()", find a relevant comparison to highlight by ploting the selected grouped data side by side.
```{r plot2}

```


## IV. Statistical tests (Ref Module 111 - Day 11 September 28. Homework 2)
Learning objectives: Perform and interpret a statistical test to compare means or proportions for 2 groups or more

1. Formulate a statistical hypothesis to test the potential relationship between your outcome and an other quantitative variable (patients'characteritics or presume risk factors) 
2. Perform the appropriate test and interpret (do not forget the potential  underlying assumption(s))
```{r tests}

```

3. Formulate a statistical hypothesis to test the potential relationship between your outcome and an other qualitative variable (patients'characteritics or presume risk factors) 
4. Perform the appropriate test and interpret (do not forget the potential  underlying assumption(s))
```{r tests}

```



## V. Linear regression (Ref. Module 106 - Day 16, November 9. Homework 3)
Learning objectives: (If appropriate) Build a linear model between 2 quantitiative variables.

1. Display the appropriate scatter plot
2. Compute the correlation coefficient 
3. Build the linear model
3. Assess the validity of the linear model using diagnostic plot
4. Interpet the model

```{r model}
```

