---
title: 'Lab 4: ANOVA'
author: "Nolwenn Le Meur"
date: "7 November 2016"
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- T
```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

#Objectives:

* Define hypothesis to be tested
* Compute and interpret statistical tests including ANOVA and Kruskall-Walli Tests

You are again interested in risk factors associated with low infant birth weight. You are given the  _birthwt_ dataset, which is composed of 189 women (rows) and 10 variables (columns). The data were collected at Baystate Medical Center, Springfield, Mass during 1986. 
```{r load data}
library(MASS)
data("birthwt")
```


## I. Infant birth weigth and _race_
You suspect a relationship between infant birth weigth and _race_

1. State the null hypothesis 
2. Choose and apply the appropriate statistical tests
3. Interpret the results

## II. Low birth weigth and _race_
You suspect a relationship between low birth weigth and _race_

1. State the null hypothesis 
2. Choose and apply the appropriate statistical test
3. Interpret the results
```{r}
fisher.test(table(birthwt$low, birthwt$race))
```

## III. Race and the others risk factors
You suspect that _race_ might be a counfonder in your analysis. Analysis the relationship between _race_ and the other risk factors.

1. State the null hypothesis 
2. Choose and apply the appropriate statistical tests
3. Interpret the results

