---
title: 'Lab 2.2: Confidence Interval'
author: "Nolwenn Le Meur & Pascal Crepey"
date: "September, 27-28 2016"
output: pdf_document
---

```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- T


```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```


Objectives:

* Interpret confidence interval of a mean or proportion
* Compute and plot confidence interval of a mean or proportion

You are interest in risk factors associated with low infant birth weight. You are given the  _birthwt_ dataset, which is composed of 189 women (rows) and 10 variables (columns). The data were collected at Baystate Medical Center, Springfield, Mass during 1986. 

1. Is it an exhaustive collection?

`r if(solution){colFmt("We have to assume that at minimum the population of interest is the population of mothers who had babies at the Baystate Medical Center, and it s very likely that there were more women than 189, it s also likely that the population of interest are mothers from Springfield, Mass. So, no, it s not exhaustive, it s a sample.", "blue")}`

2. Are we representative ? Justify.

`r if(solution){colFmt("
It actually depends on the the study population. For example, if there is not social biases in recruitment in that particular medical center we may say that our sample is representative of mothers in Springfield. If the population of Springfield is representative of the Mass. population, then it is possible that our conclusions apply for the whole state.
", "blue")}`

As the conclusion you will draw from your analysis will be at the population level, you need to describe the data in that sense. You want to get an idea of the true mean of baby's weight at birth in your population of interest.

3. What should you compute ?

`r if(solution){colFmt("
Confidence intervals for the mean weight of babies.
", "blue")}`

4. What does the code below do? Make a sentence to explicit the results.

```{r ci mean, echo=T}
library(MASS)
data(birthwt)
mean(birthwt$bwt)-1.96*sd(birthwt$bwt)/sqrt(189)
mean(birthwt$bwt)+1.96*sd(birthwt$bwt)/sqrt(189)
```

`r if(solution){colFmt("After loading the birthw data, it computes the lower and upper limits of the confidence interval at 95 where the true mean of the unknown population stands (with 5 error rate). The sd(birthwt bwt)/sqrt(189) represents the standard error (se). se is the estimated standard deviation of the unknown population.", "blue")}`

5. What happens if 1.96 is replaced by 2.57? 
`r if(solution){colFmt("
We increase the confidence level which as a consequence widen the confidence intervals.
", "blue")}`

6. Are you more confident or more precise on the results?
`r if(solution){colFmt("
I am more confident but less accurate.
", "blue")}`
7. What does the code below?
```{r pnorm, echo=T}
pnorm(1.96)
pnorm(2.57)
```

```{r ci}
# For lazy/smart users:
# install epiDisplay library for good
# install.packages("epiDisplay")
# Loading library each time you want to use it
library(epiDisplay)
# compte 95% confidence interval
ci(birthwt$bwt, alpha=0.05)
# compte 99% confidence interval
ci(birthwt$bwt, alpha=0.01)
```

Similarly, you want to assess the proportion of smoking women in your population. 

8. Using the ci() function from the _epiDisplay_ package, compute the 95% confidence interval of the proportion of smoking women in the population. Make a sentence to explicit the results.
```{r ci binomial, message=FALSE}
library(epiDisplay)
ci(birthwt$smoke)
```

You wish you were more accurate in your estimation. When "your" statistician compute the sample size, he used the code below. According to the results you selected 2 times the number of women thinking you would get a delta of 0.05. Were you right?
```{r ci binomial2, message=FALSE}
n.for.survey(0.4, delta=0.1)
```

```{r ci binomial3, message=FALSE}
n.for.survey(0.4, delta=0.05)
```

9. From the previous lab session (section 5), you know that there seem to be differences in the distribution of smoking women between races. Compute the 95% confidence interval of the proportion of smoking women by race group. Can you conclude on your hypotheses?
```{r ci binomial, message=FALSE}
library(epiDisplay)
by(birthwt$smoke, birthwt$race , ci)
```

# Homework
As for the confidence interval of the mean baby's weight, try to compute 95% CI of the proportion of smoking women in the epopulation, without the help of the ci() function but the formula from the course.
```{r ci proportion}
# get the proportion of smoking women
prop.table(table(birthwt$smoke))
0.39-1.96*sqrt((0.39*0.64)/189)
0.39+1.96*sqrt((0.39*0.64)/189)

## OR with mean(), it works here because smoke is a binary variable coded as integer in R
mean(birthwt$smoke)
mean(birthwt$smoke)+1.96*sqrt(mean(birthwt$smoke)*(1-mean(birthwt$smoke))/189)
mean(birthwt$smoke)-1.96*sqrt(mean(birthwt$smoke)*(1-mean(birthwt$smoke))/189)
```

