---
title: 'Homework3'
author: "Nolwenn Le Meur"
date: "16 December 2016"
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- F
```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

#Objectives:
* Descriptive statistics
* Graphics
* Univariate tests
* Linear regression

You are interested in air quality around Rennes area. You get data from 2001. The variables are:

- date of the measurement
- maxO3 the highest(worst) ozone level of the day
- T9, T12, T15 temperature at 9am, noon, and 3pm
- N9, N12, N15 nebulosity at 9am, noon, and 3pm
- W9, W12, W15 at 9am, noon, and 3pm
- maxO3eve the highest(worst) ozone level of the previous day 
- wind
- rain

## I. Descriptive statistics
Use the fonction load() to load the RData file as follow:

1. What are the types of the variables
```{r load data}
load("air.Rdata")
```
`r if(solution){colFmt("Quantitative continous: Date, max03, T9, T12,T15, N9, N12, N15, W9, W12, W15, max03eve; Qualitative: Wind (nominal), Rain (binary)", "blue")}`

2. Using the summary() function to compute descriptive statistical parameters. Are all the results meanfull?
```{r summary}
summary(air)
```
`r if(solution){colFmt("Date should not be summarized.", "blue")}`

3. Use the epiDisplay package and the ci() to compute the confidence
interval for the max03 variable. Explicit the result
```{r ci}
library(epiDisplay)
ci(air$maxO3)
```

## II. Graphics
4. Choose the propriate plot to represent exposition to max03 
```{r summary2}
plot(air$Date,air$maxO3, type="l", 
     xlab="Date", ylab="O3", col="red", main="Evolution of the highest level of ozone per day in Rennes in summer 2001")
```

## III. Univariate statistical tests
5. The recommended limit for ozone in the OMS 2005 Air Quality Guidelines is 100 micrograms per m3. What is the proportion of days with ozone above that threshold.
```{r split}
air$aboveT = 0
air$aboveT[air$maxO3> 100] = 1
prop.table(table(air$aboveT))
```

6. Test whether there is an association between the proportion of days with ozone above the OMS threshold and the rain 
```{r rain}
fisher.test(air$Rain, air$aboveT)
prop.table(table(air$Rain, air$aboveT), margin = 2)
```

7. Test whether there is an association between the proportion of days with ozone above the OMS threshold and the wind. 
8. In case of a relationshisp what wind seems problematic?
```{r wind}
fisher.test(air$Wind, air$aboveT)
prop.table(table(air$Wind, air$aboveT), margin = 2)
```

## IV. Regression
9. You want to model the level of ozone with the temperature and the nebulosity using a simple linear regression. Which model best explain the level of ozone? Why?  
```{r model}
regT15 = lm(maxO3~T15, data=air)
summary(regT15)
# Best R2 but
plot(regT15)
# Poor diagnostic plots
median(air$maxO3)
var.test(air$T15[air$maxO3<=81.5], air$T15[air$maxO3>81.5])
# Variance issue

# Best anwser the assumption of constant variance 
var.test(air$N15[air$maxO3<=81.5], air$N15[air$maxO3>81.5])
regN15 = lm(maxO3~N15, data=air)
summary(regN15)
# but the residuals are not normaly distributed
shapiro.test(regN15$residuals)

# it's probable that 
```

