
---
title: 'Lab: Analysis of surveillance data'
author: "Nolwenn Le Meur & Pascal Crepey"
date: "20 September 2016"
output: pdf_document
---

```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- FALSE

```


##Objectives: 
* Describe a timeserie about the surveillance of a disease.
* Use the appropriate tools to analyze it. 
* Interpret the analysis.


This lab is split in two parts. First we are going to analyze, step by step, a time serie of sales of drugs specific to scabies. Second, you'll use the same methodology to analyse influenza like illness time series in France between 2010 and 2016. 

#0. Install libraries

```{r eval=F, echo=T}
install.packages("surveillance")
install.packages("forecast")
```
```{r eval=F, echo=T}
library("surveillance")
library("forecast")
```

#1. Critical reading of a timeserie analysis
Download from Real the article "Using Google Flu Trends data in forecasting influenza-like illness related ED visits in Omaha, Nebraska" from Ozgur M. Araz, Dan Bentley, and Robert L. Muelleman (arzaz_etal.pdf).

In groups of 5, read the article but do not focus too much on the statistical details of the methods. Prepare answers to the "preliminary steps of surveillance data".


#2. Scabies
##A. Loading and formatting data 
Use the function read.csv() to load the data from the file "RawdataScabiesDrugs.csv".
Make sure that you used the right format (are you using the right field separator ?).
```{r load}
scabies<-read.csv2(file="RawdataScabiesDrugs.csv")
```

Try ?ts to get information on how to create a R timeserie. Then create two from previously loaded data: one for oral and one for topical drugs.
```{r ts, echo=T}
tsOral=ts(scabies$O, frequency = 52, start=c(2007,1))
```
```{r ts2}
tsTopical=ts(scabies$T, frequency = 52, start=c(2007,1))
```

##B. Plot data
```{r plot}
plot(scabies$O,type="l")
plot(scabies$T,type="l")
```

##C. Proceed with the preliminary steps described in the lecture
###1. Evaluate the surveillance system. What are its advantages ? What are its limits ?
###2. Describe the kind of collected data. What do you see in the data ?
###3. Assess quality of data. Is it noisy ? Do you notice trivial mistakes ? etc...

##D. Describe the data in terms of time
###Which time window is better suited ?

Try ?aggregate to change the time unit of your data. Compare the current timescale with a quaterly and a yearly timescale. Comment.

```{r aggregate}

plot(aggregate(tsTopic, nfrequency = 4), col="red")
plot(aggregate(tsOral, nfrequency = 1), col="red")

```

###Compare the oral drug sales and to topical drug sales time serie.
Just from the plot, note the differences and similarities between the two timeseries.


We do not have population or localization data associated with the time series, consequently we cannot proceed further the description.

##E. Decompose

###Use the ma() function like below to compute and plot moving averages with a window of 4 weeks, 12 weeks and 52 weeks.
```{r ma, echo=T}
plot(ma(tsOral, order = 4))
```
###Decompose the two timeseries and compare the results in terms of trends and seasonality (use decompose() and plot()). What do you notice ?

```{r decompose}
decO=decompose(tsOral)
decT=decompose(tsTopic)
```
###Try to do the same thing with stl()

```{r decomposestl}
decOs=stl(tsOral, s.window="periodic")
decTs=stl(tsTopic, s.window="periodic")
```

##F. Predict
ARIMA models are a family of models specifically designed to analyze timeseries. In case of seasonality in the timeseries, they are even able to reproduce it.
In the forecast library, the function auto.arima() will automatically select the best ARIMA model.

```{r arima}
fitO=auto.arima(tsOral, trace=T)
fitT=auto.arima(tsTopic, trace=T)
```
Once the model is selected and stored in a variable, you can then input it in the forecast function. The parameter "h" sets the number of time unit you want to forecast. Plot the results of a one year forecast for the two models and interpret the results.
```{r forecast}
plot(forecast(fitO, h=52))
plot(forecast(fitT, h=52))
```

#3. Influenza like illness (at home)
The ILI.csv file is an extraction from the French Sentinelles network website. Browse the website (https://websenti.u707.jussieu.fr/sentiweb/?lang=en) to get more information about the data (influenza like illness surveillance). 
Then follow the steps previously applied to perform an analysis of the data.

```{r loadILI}
ILI=read.csv2(file = "ILIData.csv")

plot(decompose(tsILI))

plot(stl(tsILI, s.window="periodic"))
```
