---
title: "Lab 3: Statistical tests"
author: "Nolwenn Le Meur & Pascal Crepey"
date: "27 September 2016"
output: pdf_document
---


```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- F


```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```


#Objectives:

* Define hypothesis to be tested
* Compute and interpret statistical tests of a mean
* Compute and interpret statistical tests of a proportion

You are *still* interested in risk factors associated with low infant birth weight. You are given the  _birthwt_ dataset, which is composed of 189 women (rows) and 10 variables (columns). The data were collected at Baystate Medical Center, Springfield, Mass during 1986. 

##1. Hypothesis

###1.1 We suspect that age, race, and smoking status may have a role to play in the low birth weight of newborns. Define the H0 (nil hypothesis) necessary to study the impact of each of these variables on babies's weight separately.

`r if(solution){colFmt("", "blue")}`

###1.2 Define the H1 (alternative hypothesis) necessary to study the impact of each of these variables on babies's weight separately.

`r if(solution){colFmt("", "blue")}`

##2. Performing statistical tests to compare means

###2.1 Compute 95% confidence intervals of birth weigths of babies born from smoking and non-smoking women. You can use _subset_ from the first lab to split the data.  
```{r ci} 
smoker = subset(birthwt, smoke==1)
nonsmoker = subset(birthwt, smoke==0)
ci(smoker$bwt)
ci(nonsmoker$bwt)
#or 

ci(birthwt$bwt[birthwt$smoke==1])
ci(birthwt$bwt[birthwt$smoke==0])
```

###2.2 From the previous results, do you think the statistical test will show a statistically significant results and why ?

`r if(solution){colFmt("Although the two intervals overlap, the overlaping area is very small. The probability that the true value is in this small area is small, potentially lower than 0.05.", "blue")}`

###2.3 Compute the z test seen in class by first computing m1, m2, sd1, sd2, (and n1, n2). You can use pnorm to compute your p value. $(z=(m_1 - m_2)/\sqrt{s_1^2/n_1 + s_2^2/n_2}$

```{r ztest} 
m1=mean(birthwt$bwt[birthwt$smoke==1])
m2=mean(birthwt$bwt[birthwt$smoke==0])
sd1=sd(birthwt$bwt[birthwt$smoke==1])
sd2=sd(birthwt$bwt[birthwt$smoke==0])
n1=length(birthwt$bwt[birthwt$smoke==1])
n2=length(birthwt$bwt[birthwt$smoke==0])

z=abs(m1-m2)/sqrt((sd1^2)/n1 + (sd2^2)/n2)
pnorm(z)
```

###2.4 Do the same thing but now use the t.test() function and compare the results with the z test. Look up for the conditions of applications and decide which test to use.

###2.5 Do the same thing for white mothers and black mothers (race 1 & 2), interpret and conclude.

`r if(solution){colFmt("", "blue")}`

##Homework
The Wilcoxon test will compare the two groups based on the ranks of values. It does not rely on the assumption of normality of the sample means. This assumption may not be true for small sample of very skewed individual data distribution. Redo 2.4 and 2.5 using the Wilcoxon test (wilcox.test) and compare the results you obtain with the one previously computed.

##3. Performing statistical tests to compare proportions

`r if(solution){colFmt("", "blue")}`

### 3.1 Compute the contingency table of low birth weight babies and smoking status of their mothers. Rows will have to be the smoking status and columns having (or not) a low birth weight.
```{r table}
table(birthwt$smoke, birthwt$low, dnn=c("smoking status", "low weight"))
```

###3.2 Using the Chi^2 test, compare the proportions of low birth weight babies among smoking mothers and non-smoking mothers (help("chisq.test")). Interpret the results. 
```{r chisq}
mytab=table(birthwt$smoke, birthwt$low, dnn=c("smoking status", "low weight"))
chisq.test(mytab)
```

###3.3 Would you get the same results if you presented the table with columns as rows and rows as columns ?
```{r chisq2}
# Yes
mytab2=table(birthwt$low, birthwt$smoke, dnn=c("low weight","smoking status"))
chisq.test(mytab2)
```

###3.4 I would like more precision for my p-value. I want 6 digits after the decimal point. I can only achieve this by using the X-squared computed, the number of degrees of freedom and the function _pchisq_. Please, do it.

```{r pchisq}
p=1-pchisq(q=4.2359, df=1)
```

###3.5 Compare the weight of babies according to race, comparing the 3 races at the same time. What are your conclusions ?
```{r tabrace}
tabrace=table(birthwt$race, birthwt$low, dnn=c("race", "low weight"))
chisq.test(tabrace)
```

```{r comments}
```
###3.6 Do the comparison by race two by two and compare your results with the previsous ones.

# Homework
Redo the comparisons in the third part using the fisher exact test.





