---
title: "Lab 3: Statistical tests"
author: "Nolwenn Le Meur & Pascal Cr??pey"
date: "27 September 2016"
output: pdf_document
---


```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- F


```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```


#Objectives:

* Define hypothesis to be tested
* Compute and interpret statistical tests of a mean
* Compute and interpret statistical tests of a proportion

You are interest *still* in risk factors associated with low infant birth weight. You are given the  _birthwt_ dataset, which is composed of 189 women (rows) and 10 variables (columns). The data were collected at Baystate Medical Center, Springfield, Mass during 1986. 

##1. Hypothesis

###1.1 We suspect that age, race, and smoking status may have a role to play in the low birth weight of newborns. Define the H0 (nil hypothesis) necessary to study the impact of each of these variables on babies's weight separately.

`r if(solution){colFmt("", "blue")}`

###1.2 Define the H1 (alternative hypothesis) necessary to study the impact of each of these variables on babies's weight separately.

`r if(solution){colFmt("", "blue")}`

##2. Performing statistical tests to compare means

###2.1 Compute 95% confidence intervals of birth weigths of babies born from smoking and non-smoking women. You can use _subset_ from the first lab to split the data.  
```{r ci} 
smoker = subset(birthwt, smoke==1)
nonsmoker = subset(birthwt, smoke==0)
ci(smoker$bwt)
ci(nonsmoker$bwt)
#or 

ci(birthwt$bwt[birthwt$smoke==1])
ci(birthwt$bwt[birthwt$smoke==0])
```

###2.2 From the previous results, do you think the statistical test will show a statistically significant results and why ?

`r if(solution){colFmt("Although the two intervals overlap, the overlaping area is very small. The probability that the true value is in this small area is small, potentially lower than 0.05.", "blue")}`

###2.3 Compute the z test seen in class by first computing m1, m2, sd1, sd2, (and n1, n2). You can use pnorm to compute your p value. $(z=(m_1 - m_2)/\sqrt{s_1^2/n_1 + s_2^2/n_2}$

```{r ztest} 
m1=mean(birthwt$bwt[birthwt$smoke==1])
m2=mean(birthwt$bwt[birthwt$smoke==0])
sd1=sd(birthwt$bwt[birthwt$smoke==1])
sd2=sd(birthwt$bwt[birthwt$smoke==0])
n1=length(birthwt$bwt[birthwt$smoke==1])
n2=length(birthwt$bwt[birthwt$smoke==0])

z=abs(m1-m2)/sqrt((sd1^2)/n1 + (sd2^2)/n2)
pnorm(z)
```

###2.4 Do the same thing but now use the t.test() function and compare the results with the z test. Look up for the conditions of applications and decide which test to use.

###2.5 Do the same thing for white mothers and black mothers (race 1 & 2), interpret and conclude.

`r if(solution){colFmt("", "blue")}`

```{r comments}
##3. Performing statistical tests to compare proportions

`r if(solution){colFmt("", "blue")}`


# Homework
```
