---
title: 'Lab 3: Statistical test'
author: "Nolwenn Le Meur"
date: "28 September 2016"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(eval=F, echo=F)
solution <- FALSE


```

```{r correction, eval=T, echo=F, include=F}
#Color Format
colFmt = function(x,color){
  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```


Objectives:

* Formulate statistical hypotheses
* Compare sample means and proportion
* Interpreter test results

You are interest in risk factors associated with low infant birth weight. You are given the  _birthwt_ dataset, which is composed of 189 women (rows) and 10 variables (columns). The data were collected at Baystate Medical Center, Springfield, Mass during 1986. 

1. Formulate the different hypotheses between the outcome and the different risk factors
2. What test do you use to verify each hypothese?
3 
